High-order cognition is supported by complex but compressible brain activity patterns

### 2023-03-02 paper/main.tex
checking in complete draft of discussion section

\documentclass[english, 11pt]{article}
\usepackage{graphicx}
\usepackage{amsmath}
\usepackage{hyperref}
\usepackage{setspace}
\usepackage{apacite}
\usepackage{hyperref}
\usepackage{natbib}
\usepackage{pxfonts}
\usepackage[utf8]{inputenc}
\usepackage[left=1in,right=1in,top=1in,bottom=1in]{geometry}
\usepackage[left]{lineno}
\usepackage{soul}
\linenumbers

@@

Correlations or causal associations between different brain regions
simultaneously imply that full-brain activity patterns will be compressible and
also that those activity patterns will contain redundancies. For example, the
extent that activity patterns at one brain area can be inferred or predicted
from activity patterns at other areas~\citep[e.g.,][]{OwenEtal20, ScanEtal21},
reflects overlap in the information available in or represented by those brain
areas. If brain patterns in one area are recoverable using brain patterns in
another area, then the ``signal'' used to convey the activity patterns could be
compressed by removing the recoverable activity. Predictable (and therefore
redundant) brain activity patterns are also more robust to signal corruption.
For example, even if the activity patterns at one region are unreadable or
unreliable at a given moment, that unreliability could be compensated for by
other region's activity patterns that are predictive of the unreliable region.

other themes:

- the structure of our brain patterns changes with cognitive task. maybe
something about flexibility or adaptability? why might some states benefit from
more compressibiilty or more informativeness?

- for a fixed set of "symbols" the information rate decreases as the signal is
compressed further. but our finding that each brain pattern component (symbol)
becomes more informative as cognitive richness increases suggests that the
"alphabet" of brain patterns is task dependent.

- representations themselves may also change with task demands (e.g., Mack et al, 2020)

- different brain networks appear to optimize in different ways, perhaps for different
cognitive tasks or operations.

@@

- something about the fundamental nature of brain activity and the neural basis of thought.

>>>>

\documentclass[english, 11pt]{article}
\usepackage{graphicx}
\usepackage{amsmath}
\usepackage{hyperref}
\usepackage{setspace}
\usepackage{apacite}
\usepackage{hyperref}
\usepackage[sort]{natbib}
\usepackage{pxfonts}
\usepackage[utf8]{inputenc}
\usepackage[left=1in,right=1in,top=1in,bottom=1in]{geometry}
\usepackage[left]{lineno}
\usepackage{soul}
\linenumbers

@@

Our explorations of informativeness and compressibility are related to a much
broader literature on the correlational and causal structure of brain activity
patterns~\citep{PretEtal17, OwenEtal21, RogeEtal07, RubiSpor10, SizeEtal18,
SmitEtal13b, SmitEtal13c, SrinEtal07, TomaVolk11, YeoEtal11, AdacEtal12,
BassSpor17, BullSpor09, SporHone06, SporBetz16, SporZwi04, DhamEtal08,
KorzEtal08, BrovEtal04}. Correlations or causal associations between different
brain regions simultaneously imply that full-brain activity patterns will be
compressible and also that those activity patterns will contain redundancies.
For example, the extent that activity patterns at one brain area can be
inferred or predicted from activity patterns at other
areas~\citep[e.g.,][]{OwenEtal20, ScanEtal21}, reflects overlap in the
information available in or represented by those brain areas. If brain patterns
in one area are recoverable using brain patterns in another area, then the
``signal'' used to convey the activity patterns could be compressed by removing
the recoverable activity. Predictable (and therefore redundant) brain activity
patterns are also more robust to signal corruption. For example, even if the
activity patterns at one region are unreadable or unreliable at a given moment,
that unreliability could be compensated for by other region's activity patterns
that are predictive of the unreliable region.

Our findings that informativeness and compressibility change with task demands
may follow from task-dependent changes in full-brain correlation patterns. A
number of prior studies have found that so-called ``functional connectivity''
(i.e., correlational) patterns vary reliably across tasks, events, and
situations~\citep{SimoEtal16, ColeEtal14, SmitEtal09, OwenEtal21}. By examining
how these task-dependent changes in correlations affect informativeness and
compressibility, our work suggests a potential reason for why the statistical
structure of brain activity patterns might vary with cognitive task or with
cognitive demands. For lower-level tasks, or for tasks that require relatively
little ``deep'' cognitive processing, our brains may optimize activity patterns
for robustness and redundancy over expressiveness, for example to maximize
reliability. For higher-level tasks, or for tasks that require deeper cognitive
processing, our brains may sacrifice some redundancy in favor of greater
expressiveness.

In the information theory sense~\citep{Shan48}, when a signal is transmitted
using a fixed alphabet of ``symbols,'' the information rate decreases as the
signal is compressed (e.g., fewer symbols transmitted per unit time, using an
alphabet with fewer symbols, etc.). Our finding that each individual brain
component (symbol) becomes more informative as cognitive richness increases
suggests that the ``alphabet'' of brain activity patterns is also
task-dependent. Other work suggests that the representations that are
\textit{reflected} by brain activity patterns may also change with task
demands. For example, our brains may represent the same perceptual stimulus
differently depending on which aspects of the stimulus or which combinations of
features are relevant~\citep{MackEtal20}.

Different brain networks also varied in how informative and compressible their
activity patterns were across experimental conditions (e.g.,
Fig.~\ref{fig:networks}). This might reflect evolutionary optimizations that
reflect the relevant constraints or demands placed on those networks. One
possibility is that cortex is organized in a hierarchy of networks concerned
with different levels of processing or function. To the extent that different
levels of processing (e.g., low-level sensory processing versus ``deeper''
higher-level processing) reflect different stimulus
timescales~\citep[e.g.,][]{Mann20}, the network differences we observed might
also relate to the timescales ot which each network is maximally
sensitive~\citep{RegeEtal18, BaldEtal17,LernEtal11, HassEtal08}.

@@

Cognitive neuroscientists are still grappling with basic questions about the
fundamental ``rules'' describing how our brains respond, and about how brain
activity patterns and the associated cognitive representations and computations
are linked. We identified two aspects of brain activity patterns,
informativeness and compressibiilty, that appear to change systematically with
task demands and across brain networks. Our work helps to clarify how the
``neural code'' might be structured, and how the code might vary across tasks
and brain areas.
